How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-29 tse-paper/content.tex
integrate Taico's Feedback

Despite its central role in helping developers and in enabling
novel, on-ward processing applications,
the process of how to
%automatically
% ^ we survey manual as well, though leaving it out includes too much
systematically
analyze build logs has thus far not been systematically
overviewed as a research topic, bringing us to the first
research question:
\begin{simplebox}[minipage boxed title*=-5cm]{\textbf{Research Question
1}}
What is the state-of-the-research on analyzing build logs?
\end{simplebox}

To answer this question, we start our investigation---pictured in
\Cref{fig:overview}---with an
extensive literature survey (step 1).
The survey identifies various strategies to
analyze build logs, among which are developing
custom parsers, using regular expressions, doing manual inspections,
searching for keywords, or performing natural language processing.
However, there is currently no guidance on when to use which of these
approaches.
Unfortunately, our survey also shows that
less than half of the 61 works which rely on build log analysis
describe it in sufficient detail and that
the few available implementations are all based on either a custom parser
or regular expressions.
These two approaches in particular are problematic, as their creation
and maintenance is manual and expensive, does not generalize across build
% TODO Moritz: should we say 'our' experience? or 'one authors'?
% the latter might be too cumbersome
logs from different build tools, and is, as the experience of
TravisTorrent has shown, prone to error.
In fact, some even ``strongly discourage'' the creation of such
parsers~\cite{urli2018design}.
In summary, the outcome of our literature survey shows the need for a new
generation of automated techniques to analyze build logs, which are
openly available and documented in detail, together with
guidance on when to use which technique.

@@

% turned into an itemize again cause the 'sentence' around it was
% rather broken and item-ish anyways
To fill this gap, we adapted and implemented three novel, prototypical
implementations to
automatically analyze build logs (step 2).
The ideas behind the implementations are based on promising approaches
adopted from other research fields and existing build log analysis
techniques.
In this article, we present three techniques that analyze a build log
by retrieving a substring (chunk) that describes a
specific information:
\begin{itemize}
	\item \emph{Program Synthesis by Example (PBE)}, a technique
	able to
	automatically synthesize regular expressions,
	\item a technique based on information
	retrieval, using \emph{Common Text Similarity (CTS)}
	\item inspired by the manual ad hoc search for key parts in the
	log, a
	technique employing \emph{Keyword Search (KWS)}
\end{itemize}
% a technique 1) able to automatically synthesize regular expressions,
% called \emph{Program Synthesis by Example (PBE)}, a technique
% 2) based on information
% retrieval, using \emph{Common Text Similarity (CTS)}	and 3)
% inspired by the manual ad hoc search for key parts in the log, a
% technique employing \emph{Keyword Search (KWS)}.
These three approaches---synthesized regular expressions,
text similarity,
keyword search---have
different strengths and weaknesses.
As a consequence,
in the second research question we ask:

@@

Due to this ambiguity, we develop a set of guidelines
on how to choose the most suitable build log analysis
technique for the task at hand (step 4).
We recommend PBE for use cases where the desired information is always
represented in the same structural way and high confidence in
precision and recall of the chunk retrieval is required.
The results PBE produces are suited for automatic on-ward processing.
CTS is
well-suited when the representation of the desired information varies
slightly and the output of the chunk retrieval is further processed by
a human.
In cases where the textual representation of the desired
information in the log is unpredictable or varies greatly, KWS seems
to be the best choice.
However, its low precision---it extracts a
context of multiple lines around a found keyword---makes it generally
unsuited for automatic on-ward processing.
Instead, KWS requires a human
to further inspect and interpret the output chunk.
A hybrid solution combining the three techniques might be the best
all-round solution.
Finally, we give a roadmap to guide future research in the field
of build log analysis.

@@

In this section, we give a general overview over our paper population
and then present the results of the analysis phase of the literature
mapping along RQ 1.1 through RQ 1.3.

@@

In this section, we describe what sort of information papers usually
extract from build logs, along the results depicted
in \Cref{fig:litsur_r}.

@@

\Cref{fig:litsur:use} shows the reason for the
investigated articles to analyze build logs.
The majority of articles (46\%) used the outcome of their build log
analysis to drive further research.
Seo et al.~\cite{seo2014programmers}, for example,
categorized the types of compile errors encountered at
Google.
Other articles (10\%) used the gained insights to give feedback to
developers.
For 8\% of the articles was the retrieved information the input
to a next automatic on-ward processing step:
articles chose
tests that should be part of a reduced but still effective test
suite~\cite{shi2018evaluating} or filled a data structure with
build information, an information source for further tools aiming
to fix a failing build~\cite{vassallo2018un-break}.

@@

The most mentioned method (34\%) was using a parser, where we also
included
rather imprecise statements like ``we parse the build logs''
~\cite{rahman2018impact}.
For such imprecise statements it was often not clear if they describe
a complex parser (including a tokenizer, lexer, etc.) or refer to a
simpler parsing tool based on regular expressions.
If we saw evidence for the latter, we assigned the tag
``regular expression'' instead of the tag ``parser.''
25\% of the articles described the use of regular expressions and 20\%
inspected the logs manually.
For example,
Seo et al.~\cite{seo2014programmers} developed a custom
parser to classify error messages, while Vassallo et
al.~\cite{vassallo2017a-tale} analyzed build logs with regular
expressions.
Ghaleb et al.~\cite{ghaleb2019studying} used a compound approach.
They started with manual categorizing build logs and selecting
search strings that identify the target category.
Based on the search strings they created a script that automatically
classifies the remaining logs.
Another 8\% employed machine learning such as natural language
processing or information retrieval techniques.
The rather general tags ``analysis'' and ``scan'' in
\Cref{tab:litsur:techniques}, stem---similar to the imprecise
statements about parsing---from the
words authors used to describe their techniques.
%such as
%``automatically analyzed [the build logs]''.

@@

% ----- kinds of build logs RQ1.3 -----
Several articles noted the variety in log formats for different build
tools, which requires customizing build log techniques for
every supported build tool.
% The following two sentences are passive, however flipped around
% have a different meaning (we want build tool -> study count)
\Cref{fig:litsur:log_producer} shows that only a few distinct
build tools are the target of
many studies and tools (roughly, area A).
The majority of build tools was however only targeted by one
article each (roughly, area B).
From this we follow, that the large tail of
\textbf{seldomly-investigated
environments would benefit from the
existence of a more generic solution.}
In most cases, \textbf{researchers analyze build logs from open
source projects}, obtained through Travis CI.
In addition, we saw several articles that targeted the logs of only
one specific software project.

@@

% TODO Moritz: is this okay (here)?
% maybe it's better to add it to Threats to Validity with the heading:
% \texbf{Modifying the Build Process.}
An alternative to analyzing build logs, would be to integrate plugins
that report precise data about the build process, such as the JUnit
plugin for Jenkins~\cite{jenkins2020junit-plugin}.
However, such plugins are not available for all build tools and they
do not enable the analysis of historical data---builds that ran before
the plugin was installed cannot be analyzed.
In addition to that, when using such a plugin, one would have to know
beforehand which information should be reported from a build, whereas
build log analysis enables us to analyze the wide range of information
already baked into the build logs.
Their advantage is that they can precisely report the information which
is required.

% TODO Caro: merge nicely
A downside of plugins, which are integrated into the CI flow similar to
Deflaker by Bell et al.~\cite{bell2018deflaker}, is that they are unable
to work on historic
builds, i.e., before the moment they were being deployed and that if
information has not been recorded from the beginning but turns out to
be important, it is impossible to retrieve later.

@@

% what we need is some derivation of the three techniques 1) regex 2)
%cts 3) kws, best repeated here in the discussion and then from the lit
%survey above, section 2.4.2
% TODO Caro: merge!
For this reason, we are focussing on techniques that run automatically
after being configured through examples or keywords.
\emph{Examples consist of logs and corresponding chunks.}
The user provides these examples and keywords, which are easy to adapt
whenever the user adds new examples.
A user---a developer or a researcher who wants to analyze
build logs---provides these examples and keywords, which can simply be

@@

\subsubsection{Configuration}
In- and output example pairs are the main driver of Programming by
Example; we refer to them in short as \emph{examples}.
The \emph{input} is the text of the build log file.
The \emph{output} is
a substring of the log file text, representing the
substring that the synthesized program should retrieve when
given the corresponding input file.
One or multiple examples, the
training set, \emph{configure} a specific chunk retrieval with PBE:
they define the substring of a build log that should be extracted.
The PROSE program synthesis then tries to construct a program
consistent with all training examples.
If PROSE cannot synthesize a program, e.g.,\
because the regular expression
necessary is too complicated to be synthesized, PBE returns the
error message ``no program found.''

\subsubsection{Application}
A run of PBE takes a build log file as input and applies the
synthesized regular expression program.
It then returns the substring
of the build log matched by the program.
% I hope this is fine, you can also cut this sentence if it still
% bothers you :)
If the program finds no match because the analyzed build logs
do not contain the structure the training examples define,
PBE returns an error message.

@@

To improve the similarity calculation, we prune very often or
very rarely appearing words.
Finally, the
% TODO Caro: TF-IDF undefined
algorithm weighs the vectors using TF-IDF, a best practice for natural
language queries~\cite{lee1997document}.

@@

\subsubsection{Configuration}
A set of keywords configures the chunk retrieval with KWS\@.
To better
compare KWS with PBE and CTS, we also configure it through examples.
% TODO Caro: unclear
We associate each example with keywords which appear in the targeted
chunk or close to it.

% TODO Caro: unclear.
or do you mean: same number == all searched for ?
KWS then searches for those keywords which are tied to the greatest
number of examples in the training set.
If several keywords are associated with the same number of training
examples, and no other keywords are associated with more training
examples, KWS searches for all of these keywords.

\subsubsection{Application}
For a retrieval run, we take a whole build log file as input and
search for all exact occurrences of the keywords.
As keywords are
often not directly describing the desired information, but rather
appear close to the desired information, KWS also retrieves the lines
around the found keyword.
The number of surrounding lines retrieved is
the average of lines in the specification output of the training examples.

\subsection{Random Line Retrieval (RLR)}
\label{sec:expl-rlr}

To fully comprehend how difficult the task of build log analysis really
is without any a-priori knowledge, we include a comparison with a
base line of
picking lines at random from the build log.
RLR mimics the
situation of guessing blindly which lines are interesting.
The only configuration option for RLR is the number of lines it should
retrieve.
For a fair comparison to the other techniques (whose number of
returned lines is dynamic), we configure RLR to return the average
number of lines in the chunks of the training examples.
giving it a slight advantage over guessing without previous knowledge.

@@

When we train PBE with these two examples, it produces
a regular expression program similar to the one presented in
\Cref{lst:prose-program-simplified}.
The retrieval
starts after a colon and a line separator and before a capital
letter.
It ends before two line separators,
which is consistent with the two example
chunks from \Cref{lst:chunk-example}.
This shows, that the characters structuring the text around the chunk are
crucial for regular expressions to identify a log chunk.
In fact, if we add the third example from \Cref{lst:chunk-example-3}
to the training set (which has a
different structural representation),
PBE is not able to synthesize a program anymore and returns
``\texttt{no program found}.''
We later express this difference in structural representation through
separating log chunks into \emph{structural categories}.
The chunks from \Cref{lst:chunk-example}
are in the \emph{same structural category}, while the chunk from
\Cref{lst:chunk-example-3} is in a
\emph{different structural category}.
% TODO Moritz: is this fitting here?
% or too much 'foreshadowing' too far down in the hierarchy
The results of the empirical comparison study on the chunk retrieval
techniques
show that the structural representation of the targeted chunk greatly
influences
the performance of the chunk retrieval techniques.

@@

Keywords that appear close to the chunk within
the build log configure the technique KWS\@.
For the examples from \Cref{lst:chunk-example} these could be
``FAILURE,'' ``wrong,'' or ``failed.''
The example from \Cref{lst:chunk-example-3} could be found
by searching for ``FAIL.''
If we train KWS with all three examples and the aforementioned
associated keywords, it would search for the three keywords from
\Cref{lst:chunk-example}, as they appear twice---most often within
the training set of three examples.

@@

% TODO -- this is a wrong title -- !!
% find better one
\section{Empirical Comparison Study Design}
\label{sec:study}

\begin{figure*}[tb]
	\centering
	\includegraphics[width=\textwidth, trim={1.6cm 2.5cm 0.2cm 4.8cm},
  clip]{img/study.pdf}
  % TODO wrong caption!!! find adequate one
	\caption{Design of the technique comparison study.}
	\label{fig:study}
\end{figure*}

In the literature survey we saw that few articles justify why they
use a certain build log analysis technique and concluded, that there
is a lack of guidance on when to use which technique.
% add (see \Cref{sec:lit-sur:discussion}) ?
To address this lack of guidance we conduct an empirical study
comparing the three novel chunk retrieval techniques we presented in
\Cref{sec:techniques}.
We vary the conditions under which we train and run the techniques
to answer when a technique performs best and which conditions should
influence the choice for one technique over the others.

@@

\subsection{LogChunks Data Set}
\label{sec:logchunks}
To conduct the study in this paper, we created the
\emph{LogChunks} data set~\cite{brandt2020logchunks}.
It encompasses 797 build logs from Travis CI,
stemming from a broad range of 80 GitHub repositories and 29
programming languages.
For each file, we manually ``labeled
the log part (chunk) describing why the build
failed''---the information targeted by most of the articles from our
literature survey.
We included keywords, which we
would use to search for the selected chunk within the log.
In
addition, we categorized the log chunks according to their format
within the log.
If the chunks were surrounded by the same markings
within the log, we assigned them the \emph{same structural category},
as described in \Cref{sec:crt-example}.

@@

% TODO: find better title!!!!
\subsection{Study Method}

 gives an overview over the steps within .
In the following, we explain each of the steps of our study
(see \Cref{fig:study})
and introduce the metrics
we measure to answer our research questions.

We apply each of the chunk retrieval techniques
on all of the 80 repositories represented in
\emph{LogChunks} \circlenum{1}.
For each repository, we select training sets of size 1, 2, \dots 5
examples \circlenum{2}.
We vary the training set size to measure how many examples we need to
confidently configure a technique.
The training set is so small because the examples have to be
manually created by a user.
A larger training set would oppose our goal of providing techniques
that can be configured with little effort
(see \Cref{sec:lit-sur:discussion}).

@@

\end{subfigure}\hspace{\fill}
\begin{subfigure}[tb]{\columnwidth}
		\centering
				\includegraphics[width=\columnwidth,
				clip]{img/big-study/recall-precision-categorycount-CTS.pdf}
		\caption{structural category count
		in training and test set.}
		\label{fig:recall-precision-categorycount-CTS}
\end{subfigure}
\caption{Precision, recall, and F$_{1}$-score of chunk
retrieval with Common Text Similarity (CTS) compared by \ldots}
% TODO Caro: check: where are the bean plots?
\label{fig:results-CTS}
\end{figure*}

% TODO Caro: check
% You might even say Figure 12-15 show two
% plots in the short text under the section heading.
% It just makes the
% text flow so much better.

% The following sections are each quite short...
% we could merge them, but then it loses consistency with the other
% result subsections where each plot gets an own paragraph

\Cref{fig:results-CTS} shows violin plots
of the precision,
recall, and F$_{1}$-score of the chunk retrieval runs with CTS.
The black horizontal lines show the average of these measurements.
If the figure shows no violin plot for a category, we only have one
data point or all data points have the same value.

@@

\Cref{fig:results-KWS} shows violin plots
of the precision,
recall, and F$_{1}$-score of the chunk retrieval runs with KWS.
The black horizontal lines show the average of these measurements.
If the figure shows no violin plot for a category, we only have one
data point or all data points have the same value.

@@

% M: Perhaps, above in our section about logchunks, we would need some
% basic empirical stats -- e.g., how long are the files and how much
% do we usually extract from the log
% C: yup, that would be nice
% sadly, it's quite a bit of work to get those :/
% maybe I/We'll do it at the end
\subsection{Random Line Retrieval (RLR)}
\label{sec:r:rlr}

The baseline of randomly
picking lines from build logs delivers the expected low and
``random'' results: precision ranges between 8\% and 5\%,
recall between 6\% and 8\%.
As expected, the size of the training set has no impact on precision
or recall of retrieving chunks with RLR.
With a larger number of structural categories within the training and
test sets, the precision of RLR decreases from 7\% to 0\%.
Its recall
increases from 6\%, for one structural category present, to 11\%, for
three structural categories present and drops to 0\% again for more
structural categories.

@@

\noindent
\textbf{Retrieval Output Usage.}
If the program synthesis succeeds and applying the regular expression
program yields an output, PBE has high precision and recall.
The technique
clearly identifies a failing synthesis or when the regular expression
finds no match on a build log.
Therefore, if PBE produces
an output, the user can have high confidence that it is the desired
output.
This preciseness makes output from PBE chunk retrieval
well-suited for machine consumption and therefore automatic on-ward
processing.

@@

\subsection{Recommendation of Suitable Techniques}
We saw, that each of the techniques is suited for different types of
log chunks and the output of each is suited for different uses.
Now, we unify these results into conditions to answer
our second research questions;
If researchers or developers want to retrieve chunks from build logs,
they can follow these conditions to identify the technique
suited best for their use case.
% I want to ref the figure early bc I believe the text is easier
% to follow when looking at the figure as well
We summarize these conditions in the decision tree in
\Cref{fig:crt-recommendation}.
It is built up of questions
which either lead to more questions or to a leaf node containing a
recommended technique.
At the end of this section we give two concrete examples on how to
apply our decision tree.

@@

As a final recommendation, one could create a ``super analyzer'' by
combining the different build log analysis techniques studied in this
paper.
Such a super analyzer would likely always first employ PBE
(because of its high accuracy), followed by a combination of CTS and
KWS\@.
Other than initial setup and training time, there is no
downside to this approach if it is implemented in a transparent way to
the underlying technique: the output of the super analyzer could
include from which sub-technique it originated and thus facilitate
automatic on-ward processing or interpretation of the result.

@@

% kinda uncreative..
not sure what another example would be?
% not sure if this adds much on top of the sentences further up
Overall, we believe our implementations and their weaknesses stand
prototypical for the techniques.
We tried to limit idiosyncrasies as much as possible, by choosing
standard approaches and configurations.

>>>>

Despite its central role in helping developers and in enabling
novel, onward processing applications,
the process of how to
%automatically
% ^ we survey manual as well, though leaving it out includes too much
systematically
analyze build logs has thus far not been systematically
overviewed as a research topic, bringing us to the first
research question:
% TODO Discuss: is -research fine? or state-of-the-art more common and
% still not missused (because industry m)
\begin{simplebox}[minipage boxed title*=-5cm]{\textbf{Research Question
1}}
What is the state-of-the-research on analyzing build logs?
\end{simplebox}

To answer this question, we start our investigation---pictured in
\Cref{fig:overview}---with an
extensive literature survey (step 1).
The survey identifies various strategies to
analyze build logs, among which are developing
custom parsers, using regular expressions, doing manual inspections,
searching for keywords, or performing natural language processing.
However, there is currently no guidance on when to use which of these
approaches.
Unfortunately, our survey also shows that
less than half of the 61 works which rely on build log analysis
describe it in sufficient detail and that
the few available implementations are all based on either a custom parser
or regular expressions.
These two approaches are problematic, as their creation
and maintenance is manual and expensive, does not generalize across build
% TODO Discuss/Moritz: should we say 'our' experience? or
% 'one authors'?
% the latter might be too cumbersome
logs from different build tools, and is prone to errors,
as the experience of
TravisTorrent has shown~\cite{beller2017travistorrent}.
In fact, some even ``strongly discourage'' the creation of such
parsers~\cite{urli2018design}.
In summary, the outcome of our literature survey shows the need for a new
generation of automated techniques to analyze build logs, which are
openly available and documented in detail, together with
guidance on when to use which technique.

@@

% turned into an itemize again cause the 'sentence' around it was
% rather broken and item-ish anyways
To fill this gap, we adapted and implemented three novel, prototypical
implementations to
automatically analyze build logs (step 2).
The ideas behind the implementations are based on promising approaches
adopted from other research fields and existing build log analysis
techniques.
In this article, we present three techniques that analyze a build log
by retrieving a substring (chunk) that describes a
specific information:
\begin{itemize}
	\item \emph{Program Synthesis by Example (PBE)}, a technique
	able to
	automatically synthesize regular expressions,
	\item a technique based on information
	retrieval, using \emph{Common Text Similarity (CTS)}
	\item inspired by the manual ad hoc search for key parts in the
	log, a
	technique employing \emph{Keyword Search (KWS)}
\end{itemize}
% a technique 1) able to automatically synthesize regular expressions,
% called \emph{Program Synthesis by Example (PBE)}, a technique
% 2) based on information
% retrieval, using \emph{Common Text Similarity (CTS)}	and 3)
% inspired by the manual ad hoc search for key parts in the log, a
% technique employing \emph{Keyword Search (KWS)}.
These three approaches---synthesized regular expressions,
text similarity, and
keyword search---have
different strengths and weaknesses.
As a consequence,
in the second research question we ask:

@@

Due to this ambiguity, we develop a set of guidelines
on how to choose the most suitable build log analysis
technique for the task at hand (step 4).
We recommend PBE for use cases where the desired information is always
represented in the same structural way and high confidence in
precision and recall of the chunk retrieval is required.
The results PBE produces are suited for automatic onward processing.
CTS is
well-suited when the representation of the desired information varies
slightly and the output of the chunk retrieval is further processed by
a human.
In cases where the textual representation of the desired
information in the log is unpredictable or varies greatly, KWS seems
to be the best choice.
However, its low precision---it extracts a
context of multiple lines around a found keyword---makes it generally
unsuited for automatic onward processing.
Instead, KWS requires a human
to further inspect and interpret the output chunk.
A hybrid solution combining the three techniques might be the best
all-round solution.
Finally, we give a roadmap to guide future research in the field
of build log analysis.

@@

In this section, we give a general overview over our paper population.
We then present the results of the analysis phase, along the
RQs 1.1 through 1.3.

@@

In this section, we describe what sort of information the articles in
our survey usually retrieved from build logs.
These results are depicted in \Cref{fig:litsur_r}.

@@

\Cref{fig:litsur:use} shows the reason for the
investigated articles to analyze build logs.
The majority of articles (46\%) used the outcome of their build log
analysis to drive further research.
Seo et al.~\cite{seo2014programmers}, for example,
categorized the types of compile errors encountered at
Google.
Other articles (10\%) used the gained insights to give feedback to
developers.
For 8\% of the articles was the retrieved information the input
to a next automatic onward processing step:
articles chose
tests that should be part of a reduced but still effective test
suite~\cite{shi2018evaluating} or filled a data structure with
build information, an information source for further tools aiming
to fix a failing build~\cite{vassallo2018un-break}.

@@

The most mentioned method (34\%) was using a parser, where we also
included
rather imprecise statements like ``we parse the build logs''
~\cite{rahman2018impact}.
For such imprecise statements it was often not clear if they describe
a complex parser (including a tokenizer, lexer, etc.) or refer to a
simpler parsing tool based on regular expressions.
If we saw evidence for the latter, we assigned the tag
``regular expression'' instead of the tag ``parser.''
25\% of the articles described the use of regular expressions and 20\%
inspected the logs manually.
For example,
Seo et al.~\cite{seo2014programmers} developed a custom
parser to classify error messages, while Vassallo et
al.~\cite{vassallo2017a-tale} analyzed build logs with regular
expressions.
Ghaleb et al.~\cite{ghaleb2019studying} used a compound approach.
They started with a manual categorization of build logs and selected
search strings that identify the target category.
Based on the search strings they created a script that automatically
classifies the remaining logs.
Another 8\% employed machine learning such as natural language
processing or information retrieval techniques.
The rather general tags ``analysis'' and ``scan'' in
\Cref{tab:litsur:techniques}, stem---similar to the imprecise
statements about parsing---from the
words authors used to describe their techniques.
%such as
%``automatically analyzed [the build logs]''.

@@

% ----- kinds of build logs RQ1.3 -----
Several articles noted the variety in log formats for different build
tools, which requires customizing build log techniques for
every supported build tool.
% The following two sentences are passive, however flipped around
% have a different meaning (we want build tool -> study count)
\Cref{fig:litsur:log_producer} shows that only a few distinct
build tools are the target of
many studies and tools (roughly, area A).
The majority of build tools was however only targeted by one
article each (roughly, area B).
From this we follow that the large tail of
\textbf{seldomly-investigated
environments would benefit from the
existence of a more generic solution.}
In most cases, \textbf{researchers analyze build logs from open
source projects}, obtained through Travis CI.
In addition, we saw several articles that targeted the logs of only
one specific software project.

@@

% TODO Moritz: is this okay (here)?
% maybe it's better to add it to Threats to Validity with the heading:
% \texbf{Modifying the Build Process.}
An alternative to analyzing build logs, would be to integrate plugins
that report precise data about the build process, such as the JUnit
plugin for Jenkins~\cite{jenkins2020junit-plugin} or
Deflaker by Bell et al.~\cite{bell2018deflaker}.
However, such plugins are not available for all build tools and they
do not enable the analysis of historical data---builds that ran before
the plugin was deployed cannot be analyzed.
In addition to that, when using such a plugin, one would have to know
beforehand which information should be reported from a build, whereas
build log analysis enables us to analyze the wide range of information
already baked into the build logs.
Their advantage is that they can precisely report the information which
is required.

% A downside of plugins, which are integrated into the CI flow similar to
% Deflaker by Bell et al.~\cite{bell2018deflaker}, is that they are unable
% to work on historic
% builds, i.e., before the moment they were being deployed and that if
% information has not been recorded from the beginning but turns out to
% be important, it is impossible to retrieve later.i

@@

% what we need is some derivation of the three techniques 1) regex 2)
%cts 3) kws, best repeated here in the discussion and then from the lit
%survey above, section 2.4.2
For this reason, we are focussing on techniques that run automatically
after being configured through examples or keywords.
\emph{Examples consist of logs and corresponding chunks.}
A user---a developer or a researcher who wants to analyze
build logs---provides these examples and keywords, which are easy to
adapt whenever the analysis needs to be maintained.

@@

\subsubsection{Configuration}
In- and output example pairs are the main driver of Programming by
Example; we refer to them in short as \emph{examples}.
The \emph{input} is the text of the build log file.
The \emph{output} is
a substring of the log file text, representing the
substring that the synthesized program should retrieve when
given the corresponding input file.
One or multiple examples, the
training set, \emph{configure} a specific chunk retrieval with PBE:
they define the chunk of a build log that should be retrieved.
The PROSE program synthesis then tries to construct a program
consistent with all training examples.
If PROSE cannot synthesize a program, e.g.,\
because the regular expression
necessary is too complicated to be synthesized, PBE returns the
error message ``no program found.''

\subsubsection{Application}
A run of PBE takes a build log file as input and applies the
synthesized regular expression program.
It then returns the substring
of the build log matched by the program.
% I hope this is fine, you can also cut this sentence if it still
% bothers you :)
% no match is a category later in the results section
If the program finds no match because the analyzed build log
does not contain the structure that the training examples define,
PBE returns the error message ``no match found''.

@@

To improve the similarity calculation, we prune very often or
very rarely appearing words.
Finally, the
algorithm weighs the vectors using \emph{term frequency-inverse
document frequency}, a best practice for natural
language queries~\cite{lee1997document}.

@@

\subsubsection{Configuration}
A set of keywords configures the chunk retrieval with KWS\@.
To better
compare KWS with PBE and CTS, we also configure it through examples.
We associate a list of keywords with each example.
The keywords should appear in the targeted
chunk or close to it.
Then, we merge these lists and use the most commonly occuring
keywords for the search with KWS.

\subsubsection{Application}
For a retrieval run, we take a whole build log file as input and
search for all exact occurrences of the search keywords.
As keywords are
often not directly describing the desired information, but rather
appear close to the desired information, KWS also retrieves the lines
around the found keyword.
The number of surrounding lines retrieved is
the average of lines in the specification output of the training examples.

\subsection{Random Line Retrieval (RLR)}
\label{sec:expl-rlr}
To fully comprehend how difficult the task of build log analysis really
is without any a priori knowledge, we include a comparison with a
baseline of
picking lines at random from the build log.
RLR mimics the
situation of guessing blindly which lines are interesting.
The only configuration option for RLR is the number of lines it should
retrieve.
For a fair comparison to the other techniques (whose number of
returned lines is dynamic), we configure RLR to return the average
number of lines in the chunks of the training examples.

@@

When we train PBE with these two examples, it produces
a regular expression program similar to the one presented in
\Cref{lst:prose-program-simplified}.
The retrieval
starts after a colon and a line separator and before a capital
letter.
It ends before two line separators,
which is consistent with the two example
chunks from \Cref{lst:chunk-example}.
This shows that the characters structuring the text around the chunk are
crucial for regular expressions to identify a log chunk.
In fact, if we add the third example from \Cref{lst:chunk-example-3}
to the training set (which has a
different structural representation),
PBE is not able to synthesize a program anymore and returns
``\texttt{no program found}.''
We later express this difference in structural representation through
separating log chunks into \emph{structural categories}.
The chunks from \Cref{lst:chunk-example}
are in the \emph{same structural category}, while the chunk from
\Cref{lst:chunk-example-3} is in a
\emph{different structural category}.
% TODO Moritz: is this fitting here?
% or too much 'foreshadowing' too far down in the hierarchy
The results of the empirical comparison study on the chunk retrieval
techniques
show that the structural representation of the targeted chunk greatly
influences
the performance of the chunk retrieval techniques.

@@

Keywords that appear in or close to the chunk within
the build log configure the technique KWS\@.
For the examples from \Cref{lst:chunk-example} these could be
``FAILURE,'' ``wrong,'' or ``failed.''
The example from \Cref{lst:chunk-example-3} could be found
by searching for ``FAIL.''
If we train KWS with all three examples and the aforementioned
associated keywords, it would search for the three keywords from
\Cref{lst:chunk-example}, as they appear twice---most often within
the training set of three examples.

@@

\section{Empirical Comparison Study}
\label{sec:study}

\begin{figure*}[tb]
	\centering
	\includegraphics[width=\textwidth, trim={1.6cm 2.5cm 0.2cm 4.8cm},
  clip]{img/study.pdf}
	\caption{Study design of the technique comparison study.}
	\label{fig:study}
\end{figure*}

In the literature survey we saw that few articles justify why they
use a certain build log analysis technique and concluded that there
is a lack of guidance on when to use which technique.
% add (see \Cref{sec:lit-sur:discussion}) ?
To address this lack of guidance we conduct an empirical study
comparing the three novel chunk retrieval techniques we presented in
\Cref{sec:techniques}.
We vary the conditions under which we train and run the techniques
to answer when a technique performs best and which conditions should
influence the choice for one technique over the others.

@@

\subsection{LogChunks Data Set}
\label{sec:logchunks}
To conduct the study in this paper, we created the
\emph{LogChunks} data set~\cite{brandt2020logchunks}.
It encompasses 797 build logs from Travis CI,
stemming from a broad range of 80 GitHub repositories and 29
programming languages.
For each file, we manually marked
the chunk that describes why the build
failed---the information targeted by most of the articles from our
literature survey.
To be able to configure the keyword search technique,
the data set associates each chunk with keywords that we
would use to search for the selected chunk within the log.
In
addition, we categorized the log chunks according to their format
within the log.
If the chunks were surrounded by the same markings
within the log, we assigned them the \emph{same structural category},
as described in \Cref{sec:crt-example}.

@@

\subsection{Study Design}

In the following, we explain each of the steps of our study
(see \Cref{fig:study})
and introduce the metrics that
we measure to answer our research questions.

We apply each of the chunk retrieval techniques
on all of the 80 repositories represented in
\emph{LogChunks} \circlenum{1}.
For each repository, we select training sets of size 1, 2, \dots 5
\circlenum{2}.
We vary the training set size to measure how many examples we need to
confidently configure a technique.
The training set is so small because the examples have to be
manually created by a user.
A larger training set would oppose our goal of providing techniques
that can be configured with little effort
(see \Cref{sec:lit-sur:discussion}).

@@

\end{subfigure}\hspace{\fill}
\begin{subfigure}[tb]{\columnwidth}
		\centering
				\includegraphics[width=\columnwidth,
				clip]{img/big-study/recall-precision-categorycount-CTS.pdf}
		\caption{structural category count
		in training and test set.}
		\label{fig:recall-precision-categorycount-CTS}
\end{subfigure}
\caption{Precision, recall, and F$_{1}$-score of chunk
retrieval with Common Text Similarity (CTS) compared by \ldots}
\label{fig:results-CTS}
\end{figure*}

% The following sections are each quite short...
% we could merge them, but then it loses consistency with the other
% result subsections where each plot gets an own paragraph
\Cref{fig:results-CTS} shows violin plots
of the precision,
recall, and F$_{1}$-score of the chunk retrieval runs with CTS.
The black horizontal lines show the average of these measurements.
For some measurements, the violin plot is not visible because there is
only one available data point or all data points have the same value.

@@

We present the precision,
recall, and F$_{1}$-score of the chunk retrieval runs with KWS in
\Cref{fig:results-KWS}.
Violin plots show the distribution the values of each measurement,
while the black horizontal lines show their average.
If the figure shows no violin plot for a measurement, we only have one
data point or all data points have the same value.

@@

% M: Perhaps, above in our section about logchunks, we would need some
% basic empirical stats -- e.g., how long are the files and how much
% do we usually extract from the log
% C: yup, that would be nice
% sadly, it's quite a bit of work to get those :/
% maybe I/We'll do it at the end
\subsection{Random Line Retrieval (RLR)}
\label{sec:r:rlr}
The baseline of randomly
picking lines from build logs delivers the expected low and
``random'' results: precision ranges between 5\% and 8\%,
recall between 6\% and 8\%.
As expected, the size of the training set has no impact on precision
or recall of retrieving chunks with RLR.
With a larger number of structural categories within the training and
test sets, the precision of RLR decreases from 7\% to 0\%, while
the recall varies between 0\% end 11\%.

@@

\noindent
\textbf{Retrieval Output Usage.}
If the program synthesis succeeds and applying the regular expression
program yields an output, PBE has high precision and recall.
The technique
clearly identifies a failing synthesis or when the regular expression
finds no match on a build log.
Therefore, if PBE produces
an output, the user can have high confidence that it is the desired
output.
This preciseness makes output from PBE chunk retrieval
well-suited for machine consumption and therefore automatic onward
processing.

@@

\subsection{Recommendation of Suitable Techniques}
We saw that each of the techniques is suited for different types of
log chunks and the output of each is suited for different uses.
Now, we unify these results into conditions to answer
our second research questions;
If researchers or developers want to retrieve chunks from build logs,
they can follow these conditions to identify the technique
suited best for their use case.
% I want to ref the figure early bc I believe the text is easier
% to follow when looking at the figure as well
We summarize these conditions in the decision tree in
\Cref{fig:crt-recommendation}.
It is built up of questions
which either lead to more questions or to a leaf node containing a
recommended technique.
At the end of this section we give two concrete examples on how to
apply our decision tree.

@@

As a final recommendation, one could create a ``super analyzer'' by
combining the different build log analysis techniques studied in this
paper.
Such a super analyzer would likely always first employ PBE
(because of its high accuracy), followed by a combination of CTS and
KWS\@.
Other than initial setup and training time, there is no
downside to this approach if it is implemented in a transparent way to
the underlying technique: the output of the super analyzer could
include from which sub-technique it originated and thus facilitate
automatic onward processing or interpretation of the result.

@@

% kinda uncreative..
% not sure what another example would be?
% not sure if this adds much on top of the sentences further up
Overall, we believe our implementations and their weaknesses stand
prototypical for the techniques.
We tried to limit idiosyncrasies as much as possible, by choosing
standard approaches and configurations.
